Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-22 NEUROC/sec4-experiments.tex
4.1增加了超参数对精度影响的说明

For GaAN, it is also based on the multi-head mechanism.
%
Its time complexity shoud be affected by $dim(\boldsymbol{h}^1_x)$ ($d^0_{out} = d^1_{in} = dim(\boldsymbol{h}^1_x)$), $d_a$, $d_v$, $d_m$, and the number of heads $K$.
%
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan} demonstrates that the training time increased linearly with the hyper-parameters, except for $dim(\boldsymbol{h}^1_x)$.
As $dim(\boldsymbol{h}^1_x)$ increased, the training time increased first slightly and then linearly.
We observed similar phenomena in GCN, GGNN, and GAT:
When the values of hyper-parameters were too low, GNN training could not make full use of the computing power of the GPU.
%
When the values of hyper-parameters became high enough, training time increased linearly, supporting the time complexity analysis.

@@

@@

The complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} is valid.
%
Fixing other hyper-parameters, each hyper-parameter itself affects the training time and the memory usage of a GNN Layer \emph{in a linear way}.
%
Algorithm engineers can adjust hyper-parameters according to the time complexity to avoid explosive growth in the training time and memory usage.

>>>>

For GaAN, it is also based on the multi-head mechanism.
%
Its time complexity should be affected by $dim(\boldsymbol{h}^1_x)$ ($d^0_{out} = d^1_{in} = dim(\boldsymbol{h}^1_x)$), $d_a$, $d_v$, $d_m$, and the number of heads $K$.
%
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan} demonstrates that the training time increased linearly with the hyper-parameters, except for $dim(\boldsymbol{h}^1_x)$.
As $dim(\boldsymbol{h}^1_x)$ increased, the training time increased first slightly and then linearly.
We observed similar phenomena in GCN, GGNN, and GAT:
When the values of hyper-parameters were too low, GNN training could not make full use of the computing power of the GPU.
%
When the values of hyper-parameters became high enough, training time increased linearly, supporting the time complexity analysis.

@@

% added by wangyupan, 2020-12-22 ----start----
In deep learning, accuracy is also an important indicator that algorithm engineers care about.
%
We adopted a unified standard (section \ref{sec:hyper-parameters}) to evaluate the effect of hyper-parameters on accuracy.
%
\figurename~\ref{fig:exp_hyperparameter_accuracy} shows how the accuracy varies with hyper-parameters.
%
For all models, we found that different datasets had different curve changes, that was, different datasets had different sensitivity to hyper-parameter changes.
%
This was because different datasets had different characteristics, which lead to different learning difficulties.
%

Except for the effect of the number of heads on GAT and the effect of $d_a$, $d_v$, $d_m$, the number of heads $K$ on GaAN, we found the same phenomena:
the accuracy of the model was often low when hyper-parameters were low. When a certain threshold was reached, the accuracy began to fall within the range of fluctuations.
%
We conducted additional experiments for these exceptions. Taking the effect of the number of heads on GAT as an example, 
\figurename~\ref{fig:exp_hyperparameter_accuracy_gat_small_info} shows the trend is similar to the phenomena discussed above when $d_{head}$ is adjusted to a smaller value such as 2 and 4.
%
This phenomenon showed some relationship between accuracy and model complexity.
%
When the hyper-parameter value was very small, the model complexity was often low, which limited the expressive ability of the model.
%
In this case, accuracy and model complexity were positively related.
%
However, when the hyper-parameter increased to a certain threshold, the model had sufficient learning ability, and the relationship between accuracy and model complexity was no longer certain.
% added by wangyupan, 2020-12-22 ----end----

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[width=0.25\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[width=0.25\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gat.pdf}}\\
    \subfloat[GaAN]{\includegraphics[width=0.75\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gaan.pdf}}
    \caption{Effects of hyper-parameters on accuracy of four models.}
    \label{fig:exp_hyperparameter_accuracy}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_accuracy_gat_small_info.pdf}
    \caption{Effects of the number of heads on GAT ($d_{head}$=2,4)}
    \label{fig:exp_hyperparameter_accuracy_gat_small_info}
\end{figure}

@@

The complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} was valid.
%
Fixing other hyper-parameters, each hyper-parameter itself affected the training time and the memory usage of a GNN Layer \emph{in a linear way}.
%
Algorithm engineers could adjust hyper-parameters according to the time complexity to avoid explosive growth in the training time and memory usage.
%
% added by wangyupan, 2020-12-22 ----start----
There was no definite relationship between accuracy and hyper-parameters. 
%
To obtain the optimal model, hyper-parameters needed to be manually adjusted by algorithm engineers.
% added by wangyupan, 2020-12-22 ----end----
